---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r echo=FALSE}
## Seeting global knit options
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warnings=FALSE, fig.width=10)


```


## Loading and preprocessing the data
```{r}
## List of essential packages
list <- c("dplyr", "ggplot2")
## Select those that are not installed
list <- list[!list %in% installed.packages()[, 1]]
## Install the missing packages.
if(length(list))
  lapply(list, install.packages, repos = "http://cran.us.r-project.org")

library(dplyr)
library(ggplot2)

## Open .csv file in .zip file
con <- unz("./activity.zip", "activity.csv")
## Read the data and transform into tbl_df
data <- tbl_df(read.csv(con, colClasses = c("numeric", "Date", "numeric")))
```

## What is mean total number of steps taken per day?
```{r}
## Group by date and summarise by sum without taking NA
stepPerDay <- data %>% group_by(date) %>% summarise(totalSteps = sum(steps, na.rm = TRUE))
qplot(totalSteps, data=stepPerDay, geom="histogram", 
      main="Total number of steps per day", xlab="Number of steps")

meanStep<-mean(stepPerDay$totalSteps)
medianStep<-median(stepPerDay$totalSteps)
```
The steps per day average is `r sprintf("%.1f", meanStep)` and the median is `r sprintf("%.1f", medianStep)`.

## What is the average daily activity pattern?
```{r}
## Group by interval and summarise by average without taking NA
stepPerInterval <- data %>% group_by(interval) %>% summarise(avgSteps = mean(steps, na.rm = TRUE))
qplot(stepPerInterval$interval, stepPerInterval$avgSteps, geom="line", 
      main="Average of steps across all days and 5min intervals", xlab="Interval index", ylab="Avg of steps")

intervalIndexMaxStep<-stepPerInterval$interval[which.max(stepPerInterval$avgSteps)]
```
The interval index containing the maximum number of steps is `r intervalIndexMaxStep`.

## Imputing missing values
```{r}
## Counting NA
naNumber<-sum(is.na(data$steps))
```
The number of missing values is `r naNumber`.

We will now replace the NAs found in the original data by the average on every days for the same interval calculated in the previous step.
```{r}
## Replace NA by the average on every days for this interval
naRmData <- mutate(data, steps = ifelse(is.na(steps), stepPerInterval$avgSteps, steps))

## Group by day and summarise by sum without taking NA (which is useless here since NA have been replaced)
stepPerDay <- naRmData %>% group_by(date) %>% summarise(totalSteps = sum(steps, na.rm = FALSE))
qplot(totalSteps, data=stepPerDay, geom="histogram", 
      main="Total number of steps per day (with NA replaced)", xlab="Number of steps")

meanStep1<-mean(stepPerDay$totalSteps)
medianStep1<-median(stepPerDay$totalSteps)
```
The steps per day average  (of the data with NA replaced) is `r sprintf("%.1f", meanStep1)` and the median is `r sprintf("%.1f", medianStep1)`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
## Set LOCALS to get it right in english because I'm in Belgium and speeks french
Sys.setlocale("LC_TIME", "English")

## Add weekday and weekend factor
naRmData <- mutate(naRmData, 
                   dayType = weekdays(date), 
                   dayType = as.factor(ifelse(dayType=="Sunday" | dayType=="Saturday", "weekend", "weekday")))

## Get my default LOCALS back
Sys.setlocale("LC_TIME","French_Belgium.1252")

## Group by type of day and interval
stepPerInterval<- naRmData %>% group_by(dayType, interval) %>% summarise(avgSteps = mean(steps))

qplot(interval, avgSteps, data = stepPerInterval, geom = "line", facets = dayType ~ .,
      main="Average of steps across all days and 5min intervals for weekdays and weekend", 
      xlab="Interval index", ylab="Avg of steps")
```

Generated html by knit2html("PA1_template.Rmd")